Computing in the Field of Rationals

### 2012-06-23 rational_ext.tex
_

The type safety of {{Pi}} follows directly from the corresondence
between inductive definitions of the types and the big-step
semantics. Previous work also established the following properties:

@@

We present the categorical structure with minimum commentry. More
details may be found in excellent references such as Barr and Wells
[CITE] and Selinger \cite{springerlink:10.1007/978-3-642-12821-94}.

@@

  To establish the later rigorously, we will need to show when it is
  valid to slide one wire over the other and that equivalent diagrams
  for syntactically different combinators such as
  {{(f(+)g)(;)(j(+)k)}} and {{(f(;)j)(+)(g(;)k)}} do respect
  operational equivalence.  We don't formalize the graphical notation
  in this work. Joyal et. al's work on ``planar isotopy'' and
  Selinger's survey \cite{springerlink:10.1007/978-3-642-12821-94}
  show how this has been addressed before in the categorical
  setting. In the absense of any prior knowledge of category theory
  however, our wiring diagrams may be read as the ``flow of types'' in
  a combinator-circuit.

@@

While we don't do so in this work, if we extend {{Pi}} with products
and co-products (say through the addition of information effects) it
is concievable that part of its structure collapses. This sort of
collapse, while catastrophic for algebraic structures (effectively
trivializing them), still retains some interest in computing, because
in computing we are interested in the specific operational nature (the
computational content, so to speak) of the morphisms. Anectodal evidence
follows from the fact several real-world programming language have
inconsistent type systems. blah blah blah...

@@

\paragraph*{No division by zero.} 
The important side-condition on the types not captured by the syntax
above is that there are no types of the form {{1/b}}, where
{{[^b^]=0}} (see Defn. \ref{def:size} and \ref{def:size-frac}
below). This eliminates syntaxes of the form {{1/0}}, {{1/(0+0)}},
{{1/(0*b)}} etc, from being treated as valid {{b}} types. Intuitively,
this translates into the familiar restriction that there is no
division by 0 in arithmetic or the definition of fields in abstract
algebra.

>>>>

The type safety of {{Pi}} follows directly from the correspondence
between inductive definitions of the types and the big-step
semantics. Previous work also established the following properties:

@@

We present the categorical structure with minimum commentary. More
details may be found in excellent references such as Barr and Wells
[CITE] and Selinger \cite{springerlink:10.1007/978-3-642-12821-94}.

@@

  To establish the later rigorously, we will need to show when it is
  valid to slide one wire over the other and that equivalent diagrams
  for syntactically different combinators such as
  {{(f(+)g)(;)(j(+)k)}} and {{(f(;)j)(+)(g(;)k)}} do respect
  operational equivalence.  We don't formalize the graphical notation
  in this work. Joyal et. al's work on ``planar isotopy'' and
  Selinger's survey \cite{springerlink:10.1007/978-3-642-12821-94}
  show how this has been addressed before in the categorical
  setting. In the absence of any prior knowledge of category theory
  however, our wiring diagrams may be read as the ``flow of types'' in
  a combinator-circuit.

@@

While we don't do so in this work, if we extend {{Pi}} with products
and co-products (say through the addition of information effects) it
is conceivable that part of its structure collapses. This sort of
collapse, while catastrophic for algebraic structures (effectively
trivializing them), still retains some interest in computing, because
in computing we are interested in the specific operational nature (the
computational content, so to speak) of the morphisms. Anecdotal evidence
follows from the fact several real-world programming language have
inconsistent type systems. blah blah blah...

@@

\paragraph*{No division by zero.} 
The important side-condition on the types not captured by the syntax
above is that there are no types of the form {{1/b}}, where
{{[^b^]=0}} (see Def. \ref{def:size} and \ref{def:size-frac}
below). This eliminates terms of the form {{1/0}}, {{1/(0+0)}},
{{1/(0*b)}} etc, from being treated as valid {{b}} types. Intuitively,
this translates into the familiar restriction that there is no
division by 0 in arithmetic or the definition of fields in abstract
algebra.
